### 2020-06-04 dataclean.tex
figure placement

To get some intuition for the GPU lifetimes on Titan, we give two
views of 60 randomly selected GPUs (\rev{Fig.}~\ref{fig:gpuview}) and 80
randomly selected locations (\rev{Fig.}~\ref{fig:locview}).
The GPU view
visually documents the life of each GPU unit: when it was installed and
removed at various locations, its DBE and OTB events, and the last
time it was seen. The location view documents the life of a location:
when different GPUs were installed and removed, their OTB and DBE
events, and whether a removal was the last time the unit was seen on
the system. These views were critical to understanding the data and to
verifying various data processing decisions.

The third rework cycle was used to label GPUs as {\em old} batch and
{\em new} batch. The two batches are clearly identifiable in the GPU
view of Fig.~\ref{fig:gpuview} as starting near 2017 timeframe. More
frequent OTB and DBE events are apparent in the old batch. It is also
clear from this view that practically all new GPUs stayed at their
initial install location whereas the old GPUs were occasionally
reinstalled at new locations. Nevertheless, a separate analysis
determined that the vast majority of time of the vast majority of
units is spent at one location by both the new and the old units.

>>>>

To get some intuition for the GPU lifetimes on Titan, we give two
views of \rev{80} randomly selected GPUs and \rev{80} randomly
selected locations \rev{in Fig.~\ref{fig:gpu-view}}.
The GPU view
visually documents the life of each GPU unit: when it was installed and
removed at various locations, its DBE and OTB events, and the last
time it was seen. The location view documents the life of a location:
when different GPUs were installed and removed, their OTB and DBE
events, and whether a removal was the last time the unit was seen on
the system. These views were critical to understanding the data and to
verifying various data processing decisions.

The third rework cycle was used to label GPUs as {\em old} batch and
{\em new} batch. The two batches are clearly identifiable in the
\rev{SN} view of Fig.~\ref{fig:gpu-view} as those starting near 2017
timeframe. More frequent OTB and DBE events are apparent in the old
batch. It is also clear from this view that practically all new GPUs
stayed at their initial install location whereas the old GPUs were
occasionally reinstalled at new locations. Nevertheless, a separate
analysis determined that the vast majority of time of the vast
majority of units is spent at one location by both the new and the old
units.

### 2020-06-04 tbf.tex
figure placement

\begin{figure}[bt]
  \begin{center}
    \includegraphics[trim={0 1em 0 1em},clip,width=\columnwidth]{figs/MTBF_GPUwise.pdf}
  \end{center}
  \caption{Distribution of device-level MTBFs due to DBE and OTB failures across all GPUs during the
entire lifetime of the machine. DBE and OTB events are separated for each GPU to distinguish between 
the failure pattern of each event type. Data includes both old and new batch of GPUs with the new ones 
mostly having MTBFs around or less than 3 years.}
  \label{fig:Device_MTBFs}
\end{figure}

Using the cleaned data, we analyze the inter-arrival times between the DBE and OTB events. 
This analysis is done at the device level and at the system level. 
It provides important insights into the reliability of large-scale machines, where the 
failure rate of an individual device is significantly different from the overall reliability
of the machine.

A histogram of MTBFs measured across GPUs which had at least one failure event is shown 
in \rev{Fig.}~\ref{fig:Device_MTBFs}. This is a practical assessment of device reliabilities 
as opposed to those provided in the device datasheet. The failures are tracked using the
\rev{{\em SN}} of the GPUs even though a GPU might have been placed at different locations in the 
machine during its lifetime. The time to the first failure on a device is measured by taking
the insert time as the reference point, whereas, a simple difference is taken for subsequent
failures. It can be noticed that MTBFs due to DBE and OTB failures of most GPUs are clustered around 
25,000 hours or 2.8 years. This corresponds to the lifetime of most GPUs in the system after accounting 
for relocations and replacements done in the machine. Almost all of the new batch of GPUs lie below 
the average since most have a lifetime close to 3 years. We point out
that the new batch had a much smaller number of failures than
the old batch as we highlight in the next section.

>>>>

\begin{figure*}[p]
  \begin{center}
  \includegraphics[trim=0 7 0 5,clip,width=6.5in]{figs/sample_sn.pdf} \\
  (a) \\[1em]
  \includegraphics[trim=0 7 0 5,clip,width=6.5in]{figs/sample_loc.pdf} \\
  (b)
  \end{center}
  \caption{\rev{Serial number view (a) and location view (b) of GPU
      life and failures. Both SN and locations are randomly
      selected. Legend: black dots are installs, black lines are
      lifetimes at installed location, blue squares are OTB events,
      red triangles are DBE events, and black ] are ``last seen''
      events.}}
  \label{fig:gpu-view}
\end{figure*}
Using the cleaned data, we analyze the inter-arrival times between the DBE and OTB events. 
This analysis is done at the device level and at the system level. 
It provides important insights into the reliability of large-scale machines, where the 
failure rate of an individual device is significantly different from the overall reliability
of the machine.

\begin{figure}[bt]
  \begin{center}
    \includegraphics[trim={0 1em 0 1em},clip,width=\columnwidth]{figs/MTBF_GPUwise.pdf}
  \end{center}
  \caption{Distribution of device-level MTBFs due to DBE and OTB failures across all GPUs during the
entire lifetime of the machine. DBE and OTB events are separated for each GPU to distinguish between 
the failure pattern of each event type. Data includes both old and new batch of GPUs with the new ones 
mostly having MTBFs around or less than 3 years.}
  \label{fig:Device_MTBFs}
\end{figure}
A histogram of MTBFs measured across GPUs which had at least one failure event is shown 
in \rev{Fig.}~\ref{fig:Device_MTBFs}. This is a practical assessment of device reliabilities 
as opposed to those provided in the device datasheet. The failures are tracked using the
\rev{{\em SN}} of the GPUs even though a GPU might have been placed at different locations in the 
machine during its lifetime. The time to the first failure on a device is measured by taking
the insert time as the reference point, whereas, a simple difference is taken for subsequent
failures. It can be noticed that MTBFs due to DBE and OTB failures of most GPUs are clustered around 
25,000 hours or 2.8 years. This corresponds to the lifetime of most GPUs in the system after accounting 
for relocations and replacements done in the machine. Almost all of the new batch of GPUs lie below 
the average since most have a lifetime close to 3 years. We point out
that the new batch had a much smaller number of failures than
the old batch as we highlight in the next section.
